---
title: "data-analysis"
author: "M. Gabriela Puscama"
date: "8/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.show='show')
```

```{r Load Packages}

library(tidyverse) #To import, format and wrangle data, make plots (ggplot2), and edit strings (stringr)
library(plyr) #To calculate links weights for network analysis. CAREFUL! Some functions like rename() and summarize() do not work properly on tidyverse when plyr is also loaded.

```

----------------LANGUAGE HISTORY AND DEMOGRAPHICS------------------------------
NATIVE SPEAKERS

```{r Load Data - LHQ}

data1A <- read_csv("./data/Diss_Exp1_Round1_L1Span_ListA_raw.csv")
data1B <- read_csv("./data/Diss_Exp1_Round1_L1Span_ListB_raw.csv")
data2A <- read_csv("./data/Diss_Exp1_Round2_L1Span_ListA_raw.csv")
data2B <- read_csv("./data/Diss_Exp1_Round2_L1Span_ListB_raw.csv")
data3 <- read.csv("./data/Diss_Exp1_Round3_L1Span_UniqueList_raw.csv")

```

```{r Column Formatting - LHQ}

# Add "Iteration_Survey" column to rounds 2 and 3, to standardize, using dummy ifelse statements:
data2A$Iteration_Survey <- ifelse(data2A$Gender=="Femenine", "first", "first")
data2B$Iteration_Survey <- ifelse(data2B$Gender=="Femenine", "first", "first")
data3$Iteration_Survey <- ifelse(data3$Gender=="Femenine", "first", "first")

# Select columns of interest and standardize names.
data1A_clean <- data1A %>%
  select(Iteration_Survey,
         Participant,
         Gender,
         Age,
         SpanL1,
         LangSpoken,
         'Lang_0-5',
         'Lang_6-12',
         'Lang_13-18',
         Eng_classes,
         Eng_classes_time,
         Eng_classes_amount,
         Caregivers_speak_Eng,
         Eng_travel,
         Eng_country)%>%
  rename(iteration=Iteration_Survey,
         participant=Participant,
         gender=Gender,
         age=Age,
         languages=LangSpoken,
         lang0to5='Lang_0-5',
         lang6to12='Lang_6-12',
         lang13to18='Lang_13-18',
         caregiversEng=Caregivers_speak_Eng)

data1B_clean <- data1B %>%
  select(Iteration_Survey,
         Participant,
         Gender,
         Age,
         SpanL1,
         LangSpoken,
         'Lang_0-5',
         'Lang_6-12',
         'Lang_13-18',
         Eng_classes,
         Eng_classes_time,
         Eng_classes_amount,
         Caregivers_speak_Eng,
         Eng_travel,
         Eng_country)%>%
  rename(iteration=Iteration_Survey,
         participant=Participant,
         gender=Gender,
         age=Age,
         languages=LangSpoken,
         lang0to5='Lang_0-5',
         lang6to12='Lang_6-12',
         lang13to18='Lang_13-18',
         caregiversEng=Caregivers_speak_Eng)

data2A_clean <- data2A %>%
  select(Iteration_Survey,
         Participant,
         Gender,
         Age,
         SpanL1,
         LangSpoken,
         'Lang_0-5',
         'Lang_6-12',
         'Lang_13-18',
         Eng_classes,
         Eng_classes_time,
         Eng_classes_amount,
         Caregivers_speak_Eng,
         Eng_travel,
         Eng_country)%>%
  rename(iteration=Iteration_Survey,
         participant=Participant,
         gender=Gender,
         age=Age,
         languages=LangSpoken,
         lang0to5='Lang_0-5',
         lang6to12='Lang_6-12',
         lang13to18='Lang_13-18',
         caregiversEng=Caregivers_speak_Eng)

data2B_clean <- data2B %>%
  select(Iteration_Survey,
         Participant,
         Gender,
         Age,
         SpanL1,
         LangSpoken,
         'Lang_0-5',
         'Lang_6-12',
         'Lang_13-18',
         Eng_classes,
         Eng_classes_time,
         Eng_classes_amount,
         Caregivers_speak_Eng,
         Eng_travel,
         Eng_country)%>%
  rename(iteration=Iteration_Survey,
         participant=Participant,
         gender=Gender,
         age=Age,
         languages=LangSpoken,
         lang0to5='Lang_0-5',
         lang6to12='Lang_6-12',
         lang13to18='Lang_13-18',
         caregiversEng=Caregivers_speak_Eng)

data3_clean <- data3 %>%
  select(Iteration_Survey,
         Participant,
         Gender,
         Age,
         SpanL1,
         LangSpoken,
         'Lang_0.5',
         'Lang_6.12',
         'Lang_13.18',
         Eng_classes,
         Eng_classes_time,
         Eng_classes_amount,
         Caregivers_speak_Eng,
         Eng_travel,
         Eng_country)%>%
  rename(iteration=Iteration_Survey,
         participant=Participant,
         gender=Gender,
         age=Age,
         languages=LangSpoken,
         lang0to5='Lang_0.5',
         lang6to12='Lang_6.12',
         lang13to18='Lang_13.18',
         caregiversEng=Caregivers_speak_Eng)

# Add List and Round columns using dummy ifelse statements.

data1A_clean$list <- ifelse(data1A_clean$gender=="Femenine", "A", "A")
data1B_clean$list <- ifelse(data1B_clean$gender=="Femenine", "B", "B")
data2A_clean$list <- ifelse(data2A_clean$gender=="Femenine", "A", "A")
data2B_clean$list <- ifelse(data2B_clean$gender=="Femenine", "B", "B")
data3_clean$list <- ifelse(data3_clean$gender=="Femenine", "Unique", "Unique")

data1A_clean$round <- ifelse(data1A_clean$gender=="Femenine", 1, 1)
data1B_clean$round <- ifelse(data1B_clean$gender=="Femenine", 1, 1)
data2A_clean$round <- ifelse(data2A_clean$gender=="Femenine", 2, 2)
data2B_clean$round <- ifelse(data2B_clean$gender=="Femenine", 2, 2)
data3_clean$round <- ifelse(data3_clean$gender=="Femenine", 3, 3)

```

```{r Merge Data sets - LHQ}

lhq <- rbind(data1A_clean, data1B_clean, data2A_clean, data2B_clean, data3_clean)

```

```{r Exclusions and Save - LHQ}

nrow(lhq)
#[1] 249 #One participant from round 2 did not complete the survey, that's why the initial number here is 249 instead of 250.

# Exclude participants' second iteration if they completed the survey twice, exclude those who reportedly were not raised speaking Spanish.  

repeated_survey <- length(which(lhq$iteration=="second"))
repeated_survey
#[1] 5

non_native <- length(which(lhq$SpanL1=="No"))
non_native
#[1] 3

lhq_clean <- lhq %>%
  filter(iteration == "first" & SpanL1 == "Yes")

# 1 participant did not complete the survey, 5 completed it twice (first iteration only kept), and 3 reported not to have been raised in Spanish.

nrow(lhq_clean)
#[1] 241 Final number of participants

write.csv(lhq_clean, "./data/SuppData/lhq_L1Span_clean.csv")

```

```{r Demographic and Language Tables - LHQ}

age <- summary(lhq_clean$age)
age
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#  18.00   20.00   24.00   24.29   27.00   35.00

sd(lhq_clean$age)
#[1] 4.493164

gender <- table(lhq_clean$gender)
gender
#  Female       Male Non-binary 
#  138        101          2

Eng_classes <- table(lhq_clean$Eng_classes)
100*prop.table(Eng_classes)      
#No      Yes 
#34.43983 65.56017

Eng_caregivers <- table(lhq_clean$caregiversEng)
100*prop.table(Eng_caregivers)     
#      No      Yes 
#39.83402 60.16598 

```

-----PRELIMINARY COMPARISONS BETWEEN L2 LEARNERS AND NATIVE SPEAKERS------------

Mutual Information Scores - Round 1

```{r Load Data - PMI}

pmi_Span <- read_csv("./data/pmi_Round1_L1Span.csv")
pmi_Eng <- read_csv("./data/pmi_Round1_L1Eng_prelim.csv")
links_Span <- read_csv("./data/links_L1Span1_Span.csv")
links_Eng <- read_csv("./data/links_L1Eng1_Span.csv")

```

```{r Format Columns}

pmi_Span <- pmi_Span %>%
  separate(Pair, c("source", "target"), ",", remove=FALSE) %>%
  filter(PMIClean != "NaN") %>%
  mutate(PMIClean=as.numeric(PMIClean)) %>%
  mutate(logSimpPMI = log(SimplifiedPMI), logPMI = log(PMIClean)) %>%
  mutate(logSimpPMI=as.numeric(logSimpPMI), logPMI=as.numeric(logPMI)) %>% mutate(target = str_replace_all(target, " ", "")) %>%
  mutate (target = str_replace_all(target, "_", " ")) %>%
  select(Pair:logPMI)

pmi_Eng <- pmi_Eng %>%
  separate(Pair, c("source", "target"), ",", remove=FALSE) %>%
  filter(PMIClean != "NaN") %>%
  mutate(PMIClean=as.numeric(PMIClean)) %>%
  mutate(logSimpPMI = log(SimplifiedPMI), logPMI = log(PMIClean)) %>%
  mutate(logSimpPMI=as.numeric(logSimpPMI), logPMI=as.numeric(logPMI)) %>%
  select(Pair:logPMI)

links_Span <- links_Span %>%
  select(source, target, weight)

links_Eng <- links_Eng %>%
  select(source, target, weight)

# Add L1 column, using dummy ifelse statements
pmi_Span$L1 <- ifelse(pmi_Span$PMI==0, "Span", "Span")
pmi_Eng$L1 <- ifelse(pmi_Eng$PMI==0, "Eng", "Eng")

```

```{r Add links info, Save, and Re-Open - PMI}

pmi_Span_complete <- left_join(pmi_Span, links_Span)
pmi_Eng_complete <- merge(pmi_Eng, links_Eng)

write.csv(pmi_Span_complete, "./data/pmi&links_Round1_L1Span_formatted.csv")
write.csv(pmi_Eng_complete, "./data/pmi&links_Round1_L1Eng_formatted_prelim.csv")

pmi_Span <- read_csv("./data/pmi&links_Round1_L1Span_formatted.csv")
pmi_Eng <- read_csv("./data/pmi&links_Round1_L1Eng_formatted_prelim.csv")

```

```{r Merge Datasets - PMI}

# Merge datasets
pmi_full <- rbind(pmi_Span, pmi_Eng)

```

```{r Visualizations - PMI}

pmi_plot <- pmi_full %>%
  ggplot(aes(x=logSimpPMI, fill=L1)) + geom_density(alpha=0.2) + scale_fill_manual(values=c("#000000", "#0072B2"))
pmi_plot

ggsave("./figures/pmi_plot.png", plot = last_plot())

pmi_plot2 <- pmi_full %>%
  ggplot(aes(x=logPMI, fill=L1)) + geom_density(alpha=0.2) + scale_fill_manual(values=c("#000000", "#0072B2"))
pmi_plot2

pmi_plot3 <- pmi_full %>%
  ggplot(aes(x=PMI, y=weight, fill=L1)) + geom_point() + scale_fill_manual(values=c("#000000", "#0072B2"))
pmi_plot3

```

Node Degrees - Round 1

The current dataset includes 64 L2 learners (32 per list) and 94 native speakers. To have the same number of participants on both groups, I randomly sample from the native speakers 32 participants per list, to make the comparison fair.

```{r Load Data - Degrees}

# Native speakers associations data
assoc_Span <- read_csv("./data/associations_Round1_L1Span.csv")

# L2 learners links data
links_Eng <- read_csv("./data/links_L1Eng1_Span.csv")

```

```{r Random Sample Native Speakers - Degrees}

# Clean associations data and filter Span block
assoc_Span_clean <- assoc_Span %>%
  filter (Comment != "Eng" | is.na(Comment)) %>%
  filter (Response != "NA") %>%
  filter(Block=="Span")
  
# Create separate datasets for each list
assoc_SpanA <- assoc_Span_clean %>%
  filter(List=="A")
assoc_SpanB <- assoc_Span_clean %>%
  filter(List=="B")

# Create vectors with participant IDs, to sample from
Span_participantsA <- unique(assoc_SpanA$Participant)
Span_participantsB <- unique(assoc_SpanB$Participant)

# Sample 32 native speakers per list
set.seed(20)
Span_participantsA_sample <- sample(Span_participantsA, 32, replace=FALSE)
Span_participantsB_sample <- sample(Span_participantsB, 32, replace=FALSE)

# Filter data by the sampled participants
assoc_Span_sampleA <- assoc_Span %>%
  filter(Participant %in% Span_participantsA_sample)
assoc_Span_sampleB <- assoc_Span %>%
  filter(Participant %in% Span_participantsB_sample)

# Merge sampled datasets
assoc_Span_sample <- rbind(assoc_Span_sampleA, assoc_Span_sampleB)

```

```{r Compute Link Weights - Degrees}

assoc_Span_sample <- assoc_Span_sample %>%
  rename(source=word_lemma,
         target=Response_lemma)

links_Span_sample <- ddply(assoc_Span_sample, .(source, target), summarize, weight=length(source))

```

```{r Compute Degrees and Merge Groups - Degrees}

# Compute degrees by group
degrees_Span_sample <- links_Span_sample %>%
  group_by(source) %>%
  summarize(degree_Span=length(unique(target)))

degrees_Eng <- links_Eng %>%
  select(source, target) %>%
  group_by(source) %>%
  summarize(degree_Eng=length(unique(target)))

# Merge datasets
degrees <- left_join(degrees_Span_sample, degrees_Eng)

# Filter NA rows
degrees <- degrees %>%
  filter(source != "NA")

```

```{r Visualization - Degrees}

degrees_plot <- degrees %>% 
  ggplot(aes(x=degree_Span, y=degree_Eng)) + 
  geom_point() + 
  geom_smooth(color="#44AA99") + 
  geom_abline(slope=1, intercept=0, color="#AA4499", size=1) +
  labs(title = "Node degrees by group", x = "Native Spanish speakers node degrees \n (number of different associates provided \n by the group for each prompt)", y = "L2 Spanish learners node degrees \n (number of different associates provided \n by the group for each prompt)")
degrees_plot

ggsave("./figures/degrees_plot.png", plot = last_plot())

```

Mutual Information Scores - Round 2

```{r Load Data - PMI Round 2}

pmi_Span2 <- read_csv("./data/pmi_Round2_L1Span.csv")
links_Span2 <- read_csv("./data/links_L1Span2_Span.csv")

```

```{r Format Columns Round 2}

pmi_Span2 <- pmi_Span2 %>%
  separate(Pair, c("source", "target"), ",", remove=FALSE) %>%
  filter(PMIClean != "NaN") %>%
  mutate(PMIClean=as.numeric(PMIClean)) %>%
  mutate(logSimpPMI = log(SimplifiedPMI), logPMI = log(PMIClean)) %>%
  mutate(logSimpPMI=as.numeric(logSimpPMI), logPMI=as.numeric(logPMI)) %>% mutate(target = str_replace_all(target, " ", "")) %>%
  mutate (target = str_replace_all(target, "_", " ")) %>%
  select(Pair:logPMI)

links_Span2 <- links_Span2 %>%
  select(source, target, weight)

# Add L1 column, using dummy ifelse statements
pmi_Span2$L1 <- ifelse(pmi_Span2$PMI==0, "Span", "Span")

```

```{r Add links info and Save - PMI Round 2}

pmi_Span_complete2 <- left_join(pmi_Span2, links_Span2)

write.csv(pmi_Span_complete2, "./data/pmi&links_Round2_L1Span_formatted.csv")

```

Mutual Information Scores - Round 3

```{r Load Data - PMI Round 3}

pmi_Span3 <- read_csv("./data/pmi_Round3_L1Span.csv")
links_Span3 <- read_csv("./data/links_L1Span3_Span.csv")

```

```{r Format Columns Round 3}

pmi_Span3 <- pmi_Span3 %>%
  separate(Pair, c("source", "target"), ",", remove=FALSE) %>%
  filter(PMIClean != "NaN") %>%
  mutate(PMIClean=as.numeric(PMIClean)) %>%
  mutate(logSimpPMI = log(SimplifiedPMI), logPMI = log(PMIClean)) %>%
  mutate(logSimpPMI=as.numeric(logSimpPMI), logPMI=as.numeric(logPMI)) %>% mutate(target = str_replace_all(target, " ", "")) %>%
  mutate (target = str_replace_all(target, "_", " ")) %>%
  select(Pair:logPMI)

links_Span3 <- links_Span3 %>%
  select(source, target, weight)

# Add L1 column, using dummy ifelse statements
pmi_Span3$L1 <- ifelse(pmi_Span3$PMI==0, "Span", "Span")

```

```{r Add links info and Save - PMI Round 3}

pmi_Span_complete3 <- left_join(pmi_Span3, links_Span3)

write.csv(pmi_Span_complete3, "./data/pmi&links_Round3_L1Span_formatted.csv")

```

